---
title: \vspace{-1.8cm} Relationship between transmission type and miles per gallon
output: 
  pdf_document: 
    latex_engine: xelatex
---
\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Synopsis

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- “Is an automatic or manual transmission better for MPG?”
- "Quantify the MPG difference between automatic and manual transmissions" 
    
## Exploratory Data Analysis

Data we are going to use for this document comes from the _*mtcars*_ dataset. It comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). A brief summary of data can be seen on _*Table 1*_ and _*Figure 1*_ of the Appendix.

```{r}
library(tidyverse)
library(GGally)
library(skimr)
library(corrplot)
mtcars_dataset <- mtcars
mtcars_dataset$vs <- as.factor(mtcars_dataset$vs)
mtcars_dataset$am <- as.factor(mtcars_dataset$am)
corr_matrix <- cor(select(mtcars, -c(vs, am)))
```
## Model selection

Let's start by fitting the simplest model we can regarding the given questions, that would be MPG as a function of transmission`:

```{r}
model_1 <- lm(formula = mpg~am, data = mtcars_dataset)
summary(model_1)
```
The p-value of the coefficients is below our threshold of 0.05, so there's evidence to conclude that there is some relation between the type of transmission and miles per US gallon a car can achieve. However, the model has a low R-squared and so it appears to have captured little variance, underfitting the data, resulting in low generalization of their results.

Therefore, we're going to increase the number of predictors in out linear model, and test their performance with an ANOVA test. 

To do that, let's select the new features using a correlation matrix. Recall that multicollinearity is a potential problem on regression analysis. Therefore, we're going to fit a second model with features that have low correlation (according to the *Figure 2* of the Appendix). 

```{r}
model_2 <- update(model_1, mpg~am+wt+qsec)
anova(model_1,model_2)
```

The ANOVA test shows as that our second model is significantly better than our first, mode simple model. Let's take a glance to this last model:

```{r}
summary(model_2)
```
Now, our model can explain 83% of the variability, providing a better fit. Also, looking at the plots displayed on _*Figure 3*_ of the Appendix, we can see:

* No distinct pattern in the residual vs fitted plot, thus we can infer the lack of any non-linear relationship.
* Normal distribution in the QQ plot.
* No significant outlires in the Residuals vs Leverage plot.

## Conclusion

After fitting and analyzing our model, we are capable of answering the given questions:

- “Is an automatic or manual transmission better for MPG?”

Manual transmission has proven to better while considering miles per gallon consumption. 

- "Quantify the MPG difference between automatic and manual transmissions". 

Manual transmission delivers about 2.9 miles per gallon more than automatic transmission.

\newpage

## Appendix
```{r}
skim_without_charts(mtcars_dataset)
```

```{r, fig.align="center", fig.cap="Pairs plot of dataset",out.width="75%", out.height="85%"}
ggpairs(mtcars_dataset,mapping = aes(color=am),title = "Figure 1 - Pair plot of dataset")
```

```{r, fig.align='center',fig.cap="Correlation Matrix",out.width="75%", out.height="85%"}
corrplot(corr_matrix, type = "upper",tl.col = "black",method="square",tl.srt = 45,diag = FALSE,addgrid.col = NA)
```

```{r}
par(mfrow=c(2,2))
plot(model_2)
```

